\Large \bf One-Step, Three-Factor Authentication With Custom-Fit, In-Ear EEG

### 2017-02-16 paper.tex
tightened first page

\subsection*{Abstract}
In this paper, we present a system that provides all three factors of authentication (knowledge, possession, and inherence) in a single step, using brain-based authentication via custom-fit, in-ear electroencephalography (EEG). We demonstrate collecting EEG data from seven participants using manufactured custom-fit earpieces with embedded electrodes. Across all participants, we achieve mean 0\% false acceptance and 0.017\% false rejection rates with data from only one earpiece with three electrodes. We find a 0\% false acceptance rate in a preliminary simulation of an impersonation attack. Our results indicate that in-ear EEG could provide a discreet, convenient, and useable method for one-step, three-factor authentication.

@@

We find that we can achieve low false acceptance rates (FAR) and false rejection rates (FRR) with a single, 
three-electrode earpiece. Interestingly, we found that performance improves in the ear compared to a single electrode on the scalp.
Additionally, we find that a passthoughts could not be spoofed by imposters, even when the imposters had the user's earpiece and knew their chosen passthoughts. We discuss the benefits of passthoughts for usable authentication, and raise questions about how and why passthoughts work as well as they do.

@@

By employing consumer-grade EEG (electroencephalogram), it was demonstrated in a 2013 passthoughts study that a user can submit both a knowledge factor (i.e., secret thought) and an inherence factor (i.e., brainwave signal unique to the individual) in a single step by performing a single mental task \cite{Chuang2014}. Additionally, the robustness of this method against impersonation attacks was demonstrated, including conditions where the attacker may have learned the target'۪s secret thought and/or secret task \cite{Johnson2014}.

Following this, we investigate custom-fit in-ear EEG technology as the platform for investigating the feasibility, performance, and usability of one-step three-factor authentication.
 The system we propose and test here uses the choice of a mental task or "passthought" to perform as knowledge (factor one), the uniqueness of an individual's brain activity as measured by EEG as inherence (factor two), and the physical token of custom-fit earpieces, which could easily contain a hardware key-pair, as a possession factor (factor 3).
 %% In addition to the dual knowledge and inherence factors established in previous work, this study incorporates a possession factor in the form of the EEG-sensing ear-piece(s) that are custom-fitted to and worn in their ear. These earpieces can serve as physical tokens in the same way as bank ATM cards and wearable hardware tokens by implementing a hardware key-pair.
 %% Furthermore, because the earpieces are custom-fitted to each individual, we predicted they would likely not be able to produce good electrical impedances when worn by a different individual.
%% Because three-factor authentication (3FA) requires the user to submit one distinct instance of each type of authenticator, it represents the strongest level of authentication security possible. The ability to utilize all three of these security factors in a single step by performing a mental task of a few seconds is a promising in pursuit of extremely strong security while maintaining a low amount of effort and obtrusiveness to the user.

>>>>

\subsection*{Abstract}
We present a system that provides all three factors of authentication (knowledge, possession, and inherence) in a single step, using brain-based authentication via custom-fit, in-ear electroencephalography (EEG). We demonstrate collecting EEG data from seven participants using manufactured custom-fit earpieces with embedded electrodes. Across all participants, we achieve mean 0\% false acceptance and 0.017\% false rejection rates with data from only one earpiece with three electrodes. We find a 0\% false acceptance rate in a preliminary simulation of an impersonation attack. Our results indicate that in-ear EEG could provide a discreet, convenient, and useable method for one-step, three-factor authentication.

@@

We find that we can achieve low false acceptance rates (FAR) and false rejection rates (FRR) with a single, 
three-electrode earpiece. Interestingly, we found that performance improves in the ear compared to a single electrode on the scalp.
Additionally, we find that a passthoughts could not be spoofed by imposters, even when the imposters had the user's earpiece and knew their chosen passthoughts.
We discuss future work to investigate the robustness of passthoughts in day-to-day life, and raise questions about how and why passthoughts work as well as they do.

@@

By employing consumer-grade EEG (electroencephalogram), it was demonstrated in a 2013 passthoughts study that a user can submit both a knowledge factor (i.e., secret thought) and an inherence factor (i.e., brainwave signal unique to the individual) in a single step by performing a single mental task \cite{Chuang2014}. Additionally, the robustness of this method against impersonation attacks was demonstrated, including conditions where the attacker may have learned the target'۪s secret thought \cite{Johnson2014}.

Following this work, we investigate custom-fit, in-ear EEG earpieces as a platform for one-step, three-factor authentication.
We use the choice of a mental task or "passthought" to perform as knowledge (factor one), the uniqueness of an individual's brain activity as measured by EEG as inherence (factor two), and the physical token of custom-fit earpieces, which could easily contain a hardware key-pair, as a possession factor (factor 3).
 %% In addition to the dual knowledge and inherence factors established in previous work, this study incorporates a possession factor in the form of the EEG-sensing ear-piece(s) that are custom-fitted to and worn in their ear. These earpieces can serve as physical tokens in the same way as bank ATM cards and wearable hardware tokens by implementing a hardware key-pair.
 %% Furthermore, because the earpieces are custom-fitted to each individual, we predicted they would likely not be able to produce good electrical impedances when worn by a different individual.
%% Because three-factor authentication (3FA) requires the user to submit one distinct instance of each type of authenticator, it represents the strongest level of authentication security possible. The ability to utilize all three of these security factors in a single step by performing a mental task of a few seconds is a promising in pursuit of extremely strong security while maintaining a low amount of effort and obtrusiveness to the user.
